---
title: "Calibration for iBeacons"
author: "Luis P."
date: "05/15/2015"
output: html_document
---

# The beacons

We have 6 beacons in the room:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
beacons <- data.frame(beaconNr=1:6,beaconID=c("2982-17929","11667-6916","33875-41478","33198-27248","62375-26168","55326-46713"),beaconX=c(0,-2,-2,2,3,3.4),beaconY=c(0,2.7,6.4,8.9,5.8,1.8))
```



# Calibration - Walk 1

This walk went over the beacons 1-6 and 1 again, for 10s in each of them, with 5 secons to move among beacons.

We load the JSON data measured, and we add the real values of the walk (approximated)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
require(rjson)
require(plyr)
require(plotrix)
require(hydroGOF)
require(ggplot2)

ROOMSIZE <- 9

# Load the file
lines <- paste(readLines("run1-1431699239827.json"),collapse="")
# Retouch it a bit so that it is proper json
lines <- gsub("[]", "", lines, fixed=T)
lines <- gsub("][", ",", lines, fixed=T)
# We parse the json
json_data <- fromJSON(lines)
df <- ldply(json_data, data.frame)
# We add the relative timestamp since beginning of the experiment
mintime <- min(df$timestamp)
df$reltime <- (df$timestamp - mintime)

# We add, at each point in time, the real position, and calculate the real distance from each beacon
realXseries = c(rep(0,times=10),
                seq(from=0,to=-2,length.out=5),
                rep(-2,times=10),
                seq(from=-2,to=-2,length.out=5),
                rep(-2,times=10),
                seq(from=-2,to=2,length.out=5),
                rep(2,times=10),
                seq(from=2,to=3,length.out=5),
                rep(3,times=10),
                seq(from=3,to=3.4,length.out=5),
                rep(3.4,times=10),
                seq(from=3.4,to=0,length.out=5),
                rep(0,times=10),
                0
                )

realYseries = c(rep(0,times=10),
                seq(from=0,to=2.7,length.out=5),
                rep(2.7,times=10),
                seq(from=2.7,to=6.4,length.out=5),
                rep(6.4,times=10),
                seq(from=6.4,to=8.9,length.out=5),
                rep(8.9,times=10),
                seq(from=8.9,to=5.8,length.out=5),
                rep(5.8,times=10),
                seq(from=5.8,to=1.8,length.out=5),
                rep(1.8,times=10),
                seq(from=1.8,to=0,length.out=5),
                rep(0,times=10),
                0
                )


for(i in 1:length(realXseries)){
    df[round(df$reltime/1000)==(i-1),"realX"] = realXseries[i]
    df[round(df$reltime/1000)==(i-1),"realY"] = realYseries[i]
}

for (i in 1:nrow(df)){
    df[i,"beaconX"] <- beacons[beacons$beaconID == df[i,"beaconID"],"beaconX"]
    df[i,"beaconY"] <- beacons[beacons$beaconID == df[i,"beaconID"],"beaconY"]
    df[i,"realDist"] <- sqrt((df[i,"realX"]-df[i,"beaconX"])^2 + (df[i,"realY"]-df[i,"beaconY"])^2)
}

#head(df)
```

# NEW ATTEMPTS

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(forecast)
library(caret)

beaconFits <- list()

# We calculate an improved distance for each beacon
for(i in 1:nrow(beacons)){
    beacon1 <- df[df$beaconID == beacons[i,"beaconID"],]
    beacon1[beacon1$distance > ROOMSIZE, "capdistance"] <- ROOMSIZE
    beacon1[beacon1$distance <= ROOMSIZE, "capdistance"] <- beacon1[beacon1$distance <= ROOMSIZE, "distance"]
    beacon1$madistance <- ma(ts(beacon1$capdistance),order=5)
    beacon1$maproximity <- round(ma(ts(beacon1$proximity),order=3))
    beacon1$marssi <- ma(ts(beacon1$rssi),order=5)
#    fit <- lm(realDist ~ madistance+capdistance+proximity, data=beacon1)
#    fit2 <- lm(realDist ~ madistance+maproximity, data=beacon1)
    #This one seems to give a bit better results
    fit3 <- lm(realDist ~ madistance+maproximity+marssi, data=beacon1)
    #summary(fit)
    #mean(abs(fit$residuals))
    #mean(abs(fit2$residuals))
    print(mean(abs(fit3$residuals)))
    fit4 <- train(realDist ~ madistance+maproximity+marssi, data=beacon1, method="gam")
    print(mean(abs(fit4$finalModel$residuals)))
    plot(beacon1$reltime,beacon1$distance,type="l")
    lines(beacon1$reltime,beacon1$realDist,col="red")
    #lines(beacon1$reltime,predict(fit,newdata = beacon1),col="blue")
    #lines(beacon1$reltime,predict(fit2,newdata = beacon1),col="green")
    lines(beacon1$reltime,predict(fit3,newdata = beacon1),col="orange")
    lines(beacon1$reltime,predict(fit4$finalModel,newdata = beacon1),col="green")

    df[df$beaconID == beacons[i,"beaconID"],"fittedDistLM"] <- ifelse(predict(fit3,newdata = beacon1)>0, predict(fit3,newdata = beacon1), 0) 
    df[df$beaconID == beacons[i,"beaconID"],"fittedDistGAM"] <- ifelse(predict(fit4$finalModel,newdata = beacon1)>0, predict(fit4$finalModel,newdata = beacon1), 0) 
    name <- as.character(beacons[i,"beaconID"])
    beaconFits[[name]] <- fit4$finalModel
}

head(df)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# We try trilateration with the modified distance values

pred1 <- data.frame(time=unique(round(df$reltime/1000)),realX=NA,realY=NA,predXLM=NA,predYLM=NA,predXGAM=NA,predYGAM=NA)    
for(i in unique(round(df$reltime/1000))){
    
    pred1[i,"realX"] <- unique(df[round(df$reltime/1000)==i,"realX"])[[1]]
    pred1[i,"realY"] <- unique(df[round(df$reltime/1000)==i,"realY"])[[1]]

## Plotting of the original data
#   plot(beacons$beaconX,beacons$beaconY,xlim=c(-10,11),ylim=c(-8,17))   
#    subset <- df[round(df$reltime/1000)==i,]
#    points(subset$realX,subset$realY, pch="x", cex=2)
#     for(beacon in beacons$beaconID){
#         draw.circle(beacons[beacons$beaconID==beacon,"beaconX"],beacons[beacons$beaconID==beacon,"beaconY"],radius=subset[subset$beaconID==beacon,"fittedDistLM"],border="blue")
#         draw.circle(beacons[beacons$beaconID==beacon,"beaconX"],beacons[beacons$beaconID==beacon,"beaconY"],radius=subset[subset$beaconID==beacon,"fittedDistGAM"],border="green")
#     }

    # We estimate the position, see http://gis.stackexchange.com/questions/93126/trilateration-algorithm-for-n-amount-of-points-in-r
    subset <- df[round(df$reltime/1000)==i,c("beaconID","fittedDistLM","fittedDistGAM")]
    data <- merge(subset,beacons)
    
    #This code does not work?
    #norm_vec <- function(x) sqrt(sum((x[1]-data$beaconX)^2+(x[2]-data$beaconY)^2))-sum(data$distance)
    #fit <- nlm(norm_vec,c(mean(data$beaconX),mean(data$beaconY)))
    #fit$code
    
    norm_vec <- function(x) sum(((sqrt((x[1]-data$beaconX)^2+(x[2]-data$beaconY)^2) - data$fittedDistLM)^2)/(1+data$fittedDistLM^2))
    fit <- nlm(norm_vec,c(mean(data$beaconX),mean(data$beaconY)))

    norm_vec2 <- function(x) sum(((sqrt((x[1]-data$beaconX)^2+(x[2]-data$beaconY)^2) - data$fittedDistGAM)^2)/(1+data$fittedDistGAM^2))
    fit2 <- nlm(norm_vec2,c(mean(data$beaconX),mean(data$beaconY)))

        
    pred1[i,"predXLM"] <- fit$estimate[1]
    pred1[i,"predYLM"] <- fit$estimate[2]
    pred1[i,"predXGAM"] <- fit2$estimate[1]
    pred1[i,"predYGAM"] <- fit2$estimate[2]
    
## Plotting the predicted points
#    points(fit$estimate[1],fit$estimate[2], pch="o", cex=2, col="blue")
#    points(fit2$estimate[1],fit2$estimate[2], pch="o", cex=2, col="green")
#    cat ("Press [enter] to continue")
#    line <- readline()
}
pred1$predXGAMsmooth <- ma(pred1$predXGAM, order=5)
pred1$predYGAMsmooth <- ma(pred1$predYGAM, order=5)

pred1 <- pred1[complete.cases(pred1),]
# corX <- cor(pred1$realX,pred1$predX)
# print(paste("Rsquared X=",corX^2,sep=""))
# 
# corY <- cor(pred1$realY,pred1$predY)
# print(paste("Rsquared Y=",corY^2,sep=""))
# 
# pred1$residual <- sqrt((pred1$realX-pred1$predX)^2 + (pred1$realY-pred1$predY)^2)
# print(paste("Overall RMSE=",sqrt(var(pred1$residual, na.rm=T)),sep=""))

# distanceError <- sqrt((pred1$predX - pred1$realX)^2 + (pred1$predY - pred1$realY)^2)
# mean(distanceError)
# sd(distanceError)
# distanceErrorLM <- sqrt((pred1$predXLM - pred1$realX)^2 + (pred1$predYLM - pred1$realY)^2)
# mean(distanceErrorLM)
# sd(distanceErrorLM)
distanceErrorGAM <- sqrt((pred1$predXGAM - pred1$realX)^2 + (pred1$predYGAM - pred1$realY)^2)
mean(distanceErrorGAM)
sd(distanceErrorGAM)

distanceErrorGAM2 <- sqrt((pred1$predXGAMsmooth - pred1$realX)^2 + (pred1$predYGAMsmooth - pred1$realY)^2)
mean(distanceErrorGAM2)
sd(distanceErrorGAM2)


#error <- as.data.frame(distanceError)
#ggplot(error, aes(x=distanceError)) + geom_density(alpha=.2, fill="#FF6666")
plot(pred1$time,pred1$realX,type="l",ylim=c(-5,12))
#lines(pred1$time,pred1$predXLM,col="orange")
lines(pred1$time,pred1$predXGAM,col="green")
lines(pred1$time,pred1$predXGAMsmooth,col="blue")
lines(pred1$time,pred1$realY,lty=2)
#lines(pred1$time,pred1$predXLM,col="orange")
lines(pred1$time,pred1$predYGAM,lty=2,col="green")
lines(pred1$time,pred1$predYGAMsmooth,lty=2,col="blue")

# Plot of the real and predicted space distribution
ggplot(pred1, aes(x = realX, y = realY)) + 
    geom_point(size=5) + 
    geom_point(data=pred1, mapping=aes(x=predXGAM, y=predYGAM, shape="b"),alpha=0.2,col="green",size=5) +
    geom_point(data=pred1, mapping=aes(x=predXGAMsmooth, y=predYGAMsmooth, shape="c"),alpha=0.2,col="blue",size=5)



# TODO: try weighted trilateration?


```

## Test walk (cross pattern)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load the file
lines <- paste(readLines("locator-app-1432125956044-testwalk1.json"),collapse="")
# Retouch it a bit so that it is proper json
lines <- gsub("[]", "", lines, fixed=T)
lines <- gsub("][", ",", lines, fixed=T)
# We parse the json
json_data <- fromJSON(lines)
df2 <- ldply(json_data, data.frame)
# We add the relative timestamp since beginning of the experiment
mintime <- min(df2$timestamp)
df2$reltime <- (df2$timestamp - mintime)

# We add, at each point in time, the real position, and calculate the real distance from each beacon
realXseries2 = c(rep(1,times=10),
                seq(from=1,to=1,length.out=5),
                rep(1,times=10),
                seq(from=1,to=-1,length.out=5),
                rep(-1,times=10),
                seq(from=-1,to=2,length.out=5),
                rep(2,times=10),
                seq(from=2,to=2,length.out=5),
                rep(2,times=10),
                seq(from=2,to=0,length.out=5),
                rep(0,times=10),
                0
                )

realYseries2 = c(rep(0,times=10),
                seq(from=0,to=4.5,length.out=5),
                rep(4.5,times=10),
                seq(from=4.5,to=4.5,length.out=5),
                rep(4.5,times=10),
                seq(from=4.5,to=4.5,length.out=5),
                rep(4.5,times=10),
                seq(from=4.5,to=7,length.out=5),
                rep(7,times=10),
                seq(from=7,to=7,length.out=5),
                rep(7,times=10),
                7
                )


df2$realX = NA
df2$realY = NA
for(i in 1:length(realXseries2)){
    df2[round(df2$reltime/1000)==(i-1),"realX"] = realXseries2[i]
    df2[round(df2$reltime/1000)==(i-1),"realY"] = realYseries2[i]
}


for (i in 1:nrow(df2)){
    df2[i,"beaconX"] <- beacons[beacons$beaconID == df2[i,"beaconID"],"beaconX"]
    df2[i,"beaconY"] <- beacons[beacons$beaconID == df2[i,"beaconID"],"beaconY"]
    df2[i,"realDist"] <- sqrt((df2[i,"realX"]-df2[i,"beaconX"])^2 + (df2[i,"realY"]-df2[i,"beaconY"])^2)
}


# We predict the values for this walk...
# First, we apply the trained GAM to the beacon distances
for(i in 1:nrow(beacons)){
    beacon1 <- df2[df2$beaconID == beacons[i,"beaconID"],]
    beacon1[beacon1$distance > ROOMSIZE, "capdistance"] <- ROOMSIZE
    beacon1[beacon1$distance <= ROOMSIZE, "capdistance"] <- beacon1[beacon1$distance <= ROOMSIZE, "distance"]
    beacon1$madistance <- ma(ts(beacon1$capdistance),order=5)
    beacon1$maproximity <- round(ma(ts(beacon1$proximity),order=3))
    beacon1$marssi <- ma(ts(beacon1$rssi),order=5)

    name <- as.character(beacons[i,"beaconID"])
    fit <- beaconFits[[name]]

    plot(beacon1$reltime,beacon1$distance,type="l")
    lines(beacon1$reltime,beacon1$realDist,col="red")
    lines(beacon1$reltime,predict(fit,newdata = beacon1),col="green")

    df2[df2$beaconID == beacons[i,"beaconID"],"fittedDistGAM"] <- ifelse(predict(fit,newdata = beacon1)>0, predict(fit,newdata = beacon1), 0) 
}

# Then we do the trilateration with these improved values
pred2 <- data.frame(time=unique(round(df2$reltime/1000)),realX=NA,realY=NA,predXGAM=NA,predYGAM=NA)    
for(i in unique(round(df2$reltime/1000))){
    
    pred2[i,"realX"] <- unique(df2[round(df2$reltime/1000)==i,"realX"])[[1]]
    pred2[i,"realY"] <- unique(df2[round(df2$reltime/1000)==i,"realY"])[[1]]

## Plotting of the original data
#   plot(beacons$beaconX,beacons$beaconY,xlim=c(-10,11),ylim=c(-8,17))   
#    subset <- df2[round(df2$reltime/1000)==i,]
#    points(subset$realX,subset$realY, pch="x", cex=2)
#     for(beacon in beacons$beaconID){
#         draw.circle(beacons[beacons$beaconID==beacon,"beaconX"],beacons[beacons$beaconID==beacon,"beaconY"],radius=subset[subset$beaconID==beacon,"fittedDistGAM"],border="green")
#     }

    # We estimate the position, see http://gis.stackexchange.com/questions/93126/trilateration-algorithm-for-n-amount-of-points-in-r
    subset <- df2[round(df2$reltime/1000)==i,c("beaconID","distance","fittedDistGAM")]
    data <- merge(subset,beacons)
    
    #This code does not work?
    #norm_vec <- function(x) sqrt(sum((x[1]-data$beaconX)^2+(x[2]-data$beaconY)^2))-sum(data$distance)
    #fit <- nlm(norm_vec,c(mean(data$beaconX),mean(data$beaconY)))
    #fit$code
    
    #We also try the trilateration without GAM, just in case
    norm_vec <- function(x) sum(((sqrt((x[1]-data$beaconX)^2+(x[2]-data$beaconY)^2) - data$distance)^2)/(1+data$distance^2))
    fit <- nlm(norm_vec,c(mean(data$beaconX),mean(data$beaconY)))

    norm_vec2 <- function(x) sum(((sqrt((x[1]-data$beaconX)^2+(x[2]-data$beaconY)^2) - data$fittedDistGAM)^2)/(1+data$fittedDistGAM^2))
    fit2 <- nlm(norm_vec2,c(mean(data$beaconX),mean(data$beaconY)))

    pred2[i,"predX"] <- fit$estimate[1]
    pred2[i,"predY"] <- fit$estimate[2]
        
    pred2[i,"predXGAM"] <- fit2$estimate[1]
    pred2[i,"predYGAM"] <- fit2$estimate[2]
    
## Plotting the predicted points
#    points(fit2$estimate[1],fit2$estimate[2], pch="o", cex=2, col="green")
#    cat ("Press [enter] to continue")
#    line <- readline()
}

# We try additional smoothing
pred2$predXsmooth <- ma(pred2$predX, order=5)
pred2$predYsmooth <- ma(pred2$predY, order=5)
pred2$predXGAMsmooth <- ma(pred2$predXGAM, order=5)
pred2$predYGAMsmooth <- ma(pred2$predYGAM, order=5)

pred2 <- pred2[complete.cases(pred2),]

distanceError <- sqrt((pred2$predX - pred2$realX)^2 + (pred2$predY - pred2$realY)^2)
mean(distanceError)
sd(distanceError)
distanceError2 <- sqrt((pred2$predXsmooth - pred2$realX)^2 + (pred2$predYsmooth - pred2$realY)^2)
mean(distanceError2)
sd(distanceError2)


distanceErrorGAM <- sqrt((pred2$predXGAM - pred2$realX)^2 + (pred2$predYGAM - pred2$realY)^2)
mean(distanceErrorGAM)
sd(distanceErrorGAM)
distanceErrorGAM2 <- sqrt((pred2$predXGAMsmooth - pred2$realX)^2 + (pred2$predYGAMsmooth - pred2$realY)^2)
mean(distanceErrorGAM2)
sd(distanceErrorGAM2)


#error <- as.data.frame(distanceError)
#ggplot(error, aes(x=distanceError)) + geom_density(alpha=.2, fill="#FF6666")
plot(pred2$time,pred2$realX,type="l",ylim=c(-5,12))
lines(pred2$time,pred2$predX,col="orange")
lines(pred2$time,pred2$predXsmooth,col="red")
lines(pred2$time,pred2$predXGAM,col="green")
lines(pred2$time,pred2$predXGAMsmooth,col="blue")
lines(pred2$time,pred2$realY,lty=2)
#lines(pred1$time,pred1$predXLM,col="orange")
lines(pred2$time,pred2$predY,lty=2,col="orange")
lines(pred2$time,pred2$predYsmooth,lty=2,col="red")
lines(pred2$time,pred2$predYGAM,lty=2,col="green")
lines(pred2$time,pred2$predYGAMsmooth,lty=2,col="blue")

# Plot of the real and predicted space distribution
ggplot(pred2, aes(x = realX, y = realY)) + 
    geom_point(size=5) + 
    geom_point(data=pred2, mapping=aes(x=predX, y=predY, shape="b"),alpha=0.2,col="orange",size=5) +
    geom_point(data=pred2, mapping=aes(x=predXsmooth, y=predYsmooth, shape="c"),alpha=0.2,col="red",size=5) +
    geom_point(data=pred2, mapping=aes(x=predXGAM, y=predYGAM, shape="b"),alpha=0.2,col="green",size=5) +
    geom_point(data=pred2, mapping=aes(x=predXGAMsmooth, y=predYGAMsmooth, shape="c"),alpha=0.2,col="blue",size=5)

```

* As we can see, error is huge (2.4m with the GAM-trained trilateration)! 
* Smoothing of the predicted coordinates seems to help a bit in every case (2.4m vs. 2.7m)
* And actually, it works better if we use the (smoothed) plain trilateration, not the trilateration trained through GAM in a calibration walk (1.5m vs. 2.4m)!


-------------

# OLD ATTEMPTS

## Model 1: per-sample nonlinear trilateration (no training)


```{r}

#First attempt... non linear model trilateration, not training
pred1 <- data.frame(time=unique(round(df$reltime/1000)),realX=NA,realY=NA,predX=NA,predY=NA)    
for(i in unique(round(df$reltime/1000))){
    
    pred1[i,"realX"] <- unique(df[round(df$reltime/1000)==i,"realX"])[[1]]
    pred1[i,"realY"] <- unique(df[round(df$reltime/1000)==i,"realY"])[[1]]

## Plotting of the original data
#   plot(beacons$beaconX,beacons$beaconY,xlim=c(-10,11),ylim=c(-8,17))   
#    subset <- df[round(df/1000)==i,]
#    points(subset$realX,subset$realY, pch="x", cex=2)
#     for(beacon in beacons$beaconID){
#         draw.circle(beacons[beacons$beaconID==beacon,"beaconX"],beacons[beacons$beaconID==beacon,"beaconY"],radius=subset[subset$beaconID==beacon,"distance"],border="blue")
#     }

    # We estimate the position, see http://gis.stackexchange.com/questions/93126/trilateration-algorithm-for-n-amount-of-points-in-r
    subset <- df[round(df$reltime/1000)==i,c("beaconID","distance")]
    data <- merge(subset,beacons)
    
    #This code does not work?
    #norm_vec <- function(x) sqrt(sum((x[1]-data$beaconX)^2+(x[2]-data$beaconY)^2))-sum(data$distance)
    #fit <- nlm(norm_vec,c(mean(data$beaconX),mean(data$beaconY)))
    #fit$code
    
    norm_vec <- function(x) sum(((sqrt((x[1]-data$beaconX)^2+(x[2]-data$beaconY)^2) - data$distance)^2)/(1+data$distance^2))
    fit <- nlm(norm_vec,c(mean(data$beaconX),mean(data$beaconY)))
    
    pred1[i,"predX"] <- fit$estimate[1]
    pred1[i,"predY"] <- fit$estimate[2]
    
## Plotting the predicted points
#   points(fit$estimate[1],fit$estimate[2], pch="o", cex=2)
#   cat ("Press [enter] to continue")
#   line <- readline()
}

pred1 <- pred1[complete.cases(pred1),]
corX <- cor(pred1$realX,pred1$predX)
print(paste("Rsquared X=",corX^2,sep=""))

corY <- cor(pred1$realY,pred1$predY)
print(paste("Rsquared Y=",corY^2,sep=""))

pred1$residual <- sqrt((pred1$realX-pred1$predX)^2 + (pred1$realY-pred1$predY)^2)
print(paste("Overall RMSE=",sqrt(var(pred1$residual, na.rm=T)),sep=""))

distanceError <- sqrt((pred1$predX - pred1$realX)^2 + (pred1$predY - pred1$realY)^2)
mean(distanceError)
sd(distanceError)

error <- as.data.frame(distanceError)
ggplot(error, aes(x=distanceError)) + geom_density(alpha=.2, fill="#FF6666")

```

## Model 2: generalized linear model training

```{r}

# Second attempt training with generalized linear model
# We reformat our data to have more columns
pred2 <- data.frame(time=unique(round(df$reltime/1000)),realX=NA,realY=NA)
for(t in pred2$time){
    pred2[pred2$time==t,"realX"] <- unique(df[round(df$reltime/1000)==t,"realX"])[[1]]
    pred2[pred2$time==t,"realY"] <- unique(df[round(df$reltime/1000)==t,"realY"])[[1]]
}
for(beacon in beacons$beaconID){
    label = beacons[beacons$beaconID==beacon,"beaconNr"];
    # We create the field distance
    pred2[,paste("distance",label,sep="")] <- NA
    # We go though the time values and fill in the data
    for(t in pred2$time){
        pred2[pred2$time==t,paste("distance",label,sep="")] <- df[round(df$reltime/1000)==t & df$beaconID==beacon,"distance"][1]
    }

    # We create the field rssi
    pred2[,paste("rssi",label,sep="")] <- NA
    # We go though the time values and fill in the data
    for(t in pred2$time){
        pred2[pred2$time==t,paste("rssi",label,sep="")] <- df[round(df$reltime/1000)==t & df$beaconID==beacon,"rssi"][1]
    }

    # We create the field proximity
    pred2[,paste("proximity",label,sep="")] <- NA
    # We go though the time values and fill in the data
    for(t in pred2$time){
        pred2[pred2$time==t,paste("proximity",label,sep="")] <- df[round(df$reltime/1000)==t & df$beaconID==beacon,"proximity"][1]
    }

}

fitlm <- lm(cbind(realX,realY) ~ . - time, data=pred2)
summary(fitlm)

#predictions <- predict(fitlm, data.frame=pred2[,-c("realX","realY")])
#We plot the predictions
# count=0
# for(i in pred2$time){
#     count <- count+1
#     plot(beacons$beaconX,beacons$beaconY,xlim=c(-10,11),ylim=c(-8,17))
#     
#     subset <- df[round(df$reltime/1000)==i,]
#     points(subset$realX,subset$realY, pch="x", cex=2)
#     for(beacon in beacons$beaconID){
#         draw.circle(beacons[beacons$beaconID==beacon,"beaconX"],beacons[beacons$beaconID==beacon,"beaconY"],radius=subset[subset$beaconID==beacon,"distance"],border="blue")
#     }
#     points(predictions[count,"realX"],predictions[count,"realY"], pch="o", cex=2)
#     cat ("Press [enter] to continue")
#     line <- readline()
# }

predictions <- predict(fitlm,newdata=pred2)
pred2$predX <- predictions[,1]
pred2$predY <- predictions[,2]
pred2 <- pred2[complete.cases(pred2),]
distanceError <- sqrt((pred2$predX - pred2$realX)^2 + (pred2$predY - pred2$realY)^2)
mean(distanceError)
sd(distanceError)

```

As we see from the R-squared, the GLM training performs a bit better (in the training dataset!)

## Model 3: linear model based on time-smoothing of distances to beacon only

```{r}
# TODO: Third attempt training with smoothing the distances only
pred3 <- pred2

for (i in 1:6){
    column <- pred3[,paste("distance",i,sep="")]

    #We do filtering before smoothing the signal, taking into account the size of the room (9m longer diameter)
    column[column>ROOMSIZE]<-ROOMSIZE
    
    ld <- loess(column~time,pred3,span=0.15)
    pred3[,paste("loessdistance",i,sep="")] <- predict(ld,pred3)
#    plot(pred3$time,pred3[,paste("distance",i,sep="")], type="l")
#    lines(pred3$time,pred3[,paste("loessdistance",i,sep="")],col=3)
}

fitloesslm <- lm(cbind(realX,realY) ~ loessdistance1+loessdistance2+loessdistance3+loessdistance4+loessdistance5+loessdistance6, data=pred3)
summary(fitloesslm)

```

... and a similar variant taking into acount the interactions

```{r}
fitloesslm2 <- lm(cbind(realX,realY) ~ loessdistance1*loessdistance2*loessdistance3*loessdistance4*loessdistance5*loessdistance6, data=pred3)
summary(fitloesslm2)

```

This last one performs fantastically better (on the training dataset!). Let's plot it:

```{r}
predictions <- predict(fitloesslm2, newdata=pred3)
# #We plot the predictions
# count=0
# for(i in pred3$time){
#     count <- count+1
#     plot(beacons$beaconX,beacons$beaconY,xlim=c(-10,11),ylim=c(-8,17))
#     
#     subset <- df[round(df$reltime/1000)==i,]
#     points(subset$realX,subset$realY, pch="x", cex=2)
#     for(beacon in beacons$beaconID){
#         draw.circle(beacons[beacons$beaconID==beacon,"beaconX"],beacons[beacons$beaconID==beacon,"beaconY"],radius=subset[subset$beaconID==beacon,"distance"],border="blue")
#     }
#     points(predictions[count,"realX"],predictions[count,"realY"], pch="o", cex=2)
#     cat ("Press [enter] to continue")
#     line <- readline()
# }

pred3$predX <- predictions[,1]
pred3$predY <- predictions[,2]

distanceError <- sqrt((pred3$predX - pred3$realX)^2 + (pred3$predY - pred3$realY)^2)
mean(distanceError)
sd(distanceError)

```

We see it is **REALLY** close to the mark at all times, within 0.5/1 meter of the real position. But... is it overfitting (to a walk around the walls)? Let's try different walks:

## Model 4: per-sample nonlinear trilateration with smoothing


```{r}

#First attempt... non linear model trilateration, not training
pred4 <- data.frame(time=unique(round(df$reltime/1000)),realX=NA,realY=NA,predX=NA,predY=NA)    
for(i in unique(round(df$reltime/1000))){
    
    pred4[i,"realX"] <- unique(df[round(df$reltime/1000)==i,"realX"])[[1]]
    pred4[i,"realY"] <- unique(df[round(df$reltime/1000)==i,"realY"])[[1]]

## Plotting of the original data
#   plot(beacons$beaconX,beacons$beaconY,xlim=c(-10,11),ylim=c(-8,17))   
#    subset <- df[round(df/1000)==i,]
#    points(subset$realX,subset$realY, pch="x", cex=2)
#     for(beacon in beacons$beaconID){
#         draw.circle(beacons[beacons$beaconID==beacon,"beaconX"],beacons[beacons$beaconID==beacon,"beaconY"],radius=subset[subset$beaconID==beacon,"distance"],border="blue")
#     }

    # We estimate the position, see http://gis.stackexchange.com/questions/93126/trilateration-algorithm-for-n-amount-of-points-in-r
    subset <- df[round(df$reltime/1000)==i,c("beaconID","distance")]
    data <- merge(subset,beacons)
    
    #This code does not work?
    #norm_vec <- function(x) sqrt(sum((x[1]-data$beaconX)^2+(x[2]-data$beaconY)^2))-sum(data$distance)
    #fit <- nlm(norm_vec,c(mean(data$beaconX),mean(data$beaconY)))
    #fit$code
    
    norm_vec <- function(x) sum(((sqrt((x[1]-data$beaconX)^2+(x[2]-data$beaconY)^2) - data$distance)^2)/(1+data$distance^2))
    fit <- nlm(norm_vec,c(mean(data$beaconX),mean(data$beaconY)))
    
    pred4[i,"predX"] <- fit$estimate[1]
    pred4[i,"predY"] <- fit$estimate[2]
    
## Plotting the predicted points
#   points(fit$estimate[1],fit$estimate[2], pch="o", cex=2)
#   cat ("Press [enter] to continue")
#   line <- readline()
}

pred4 <- pred4[complete.cases(pred4),]

smoothfitX <- loess(predX~time,pred4,span=0.15)
smoothfitY <- loess(predY~time,pred4,span=0.15)
pred4$smpredX <- predict(smoothfitX,pred4)
pred4$smpredY <- predict(smoothfitY,pred4)

ggplot(pred4, aes(x = time, y = realX)) + geom_line(col="blue",size=1) + geom_line(data=pred4, aes(x=time,y=predX),col="black",size=1) + geom_line(data=pred4, aes(x=time,y=smpredX),col="red",size=1)

ggplot(pred4, aes(x = time, y = realY)) + geom_line(col="blue",size=1) + geom_line(data=pred4, aes(x=time,y=predY),col="black",size=1) + geom_line(data=pred4, aes(x=time,y=smpredY),col="red",size=1)


corX <- cor(pred4$realX,pred4$predX)
print(paste("Rsquared X=",corX^2,sep=""))

corY <- cor(pred4$realY,pred4$predY)
print(paste("Rsquared Y=",corY^2,sep=""))

pred4$residual <- sqrt((pred4$realX-pred4$predX)^2 + (pred4$realY-pred4$predY)^2)
print(paste("Overall RMSE=",sqrt(var(pred4$residual, na.rm=T)),sep=""))

distanceError <- sqrt((pred4$predX - pred4$realX)^2 + (pred4$predY - pred4$realY)^2)
mean(distanceError)
sd(distanceError)

error <- as.data.frame(distanceError)
ggplot(error, aes(x=distanceError)) + geom_density(alpha=.2, fill="#FF6666")
```


## Model 5: per-sample nonlinear trilateration with smoothing BEFORE


```{r}
####### REVIEW!!!!!!!!!!!!!

# We smooth the initial distance signal for each beacon
dfsmooth <- data.frame()
for(i in beacons$beaconID){
    
    subset <- df[df$beaconID==i,]
    
    #p <- ggplot(subset, aes(x=reltime,y=distance))+geom_line(col="blue",size=1)
    
    smoothfit <- loess(distance~reltime,subset,span=0.15)
    subset$newdistance <- predict(smoothfit,subset)
    
    #print(p + geom_line(data=subset, aes(x=reltime,y=newdistance), col="red",size=1))
    
    #print("more?")
    #line <- readline()
    if(nrow(dfsmooth)==0) dfsmooth <- subset
    else dfsmooth <- rbind(dfsmooth,subset)
}


#First attempt... non linear model trilateration, not training
pred4 <- data.frame(time=unique(round(dfsmooth$reltime/1000)),realX=NA,realY=NA,predX=NA,predY=NA)    
for(i in unique(round(dfsmooth$reltime/1000))){
    
    pred4[i,"realX"] <- unique(dfsmooth[round(dfsmooth$reltime/1000)==i,"realX"])[[1]]
    pred4[i,"realY"] <- unique(dfsmooth[round(dfsmooth$reltime/1000)==i,"realY"])[[1]]

## Plotting of the original data
#   plot(beacons$beaconX,beacons$beaconY,xlim=c(-10,11),ylim=c(-8,17))   
#    subset <- df[round(df/1000)==i,]
#    points(subset$realX,subset$realY, pch="x", cex=2)
#     for(beacon in beacons$beaconID){
#         draw.circle(beacons[beacons$beaconID==beacon,"beaconX"],beacons[beacons$beaconID==beacon,"beaconY"],radius=subset[subset$beaconID==beacon,"distance"],border="blue")
#     }

    # We estimate the position, see http://gis.stackexchange.com/questions/93126/trilateration-algorithm-for-n-amount-of-points-in-r
    subset <- dfsmooth[round(dfsmooth$reltime/1000)==i,c("beaconID","distance")]
    data <- merge(subset,beacons)
    
    #This code does not work?
    #norm_vec <- function(x) sqrt(sum((x[1]-data$beaconX)^2+(x[2]-data$beaconY)^2))-sum(data$distance)
    #fit <- nlm(norm_vec,c(mean(data$beaconX),mean(data$beaconY)))
    #fit$code
    
    norm_vec <- function(x) sum(((sqrt((x[1]-data$beaconX)^2+(x[2]-data$beaconY)^2) - data$distance)^2)/(1+data$newdistance^2))
    fit <- nlm(norm_vec,c(mean(data$beaconX),mean(data$beaconY)))
    
    pred4[i,"predX"] <- fit$estimate[1]
    pred4[i,"predY"] <- fit$estimate[2]
    
## Plotting the predicted points
#   points(fit$estimate[1],fit$estimate[2], pch="o", cex=2)
#   cat ("Press [enter] to continue")
#   line <- readline()
}

pred4 <- pred4[complete.cases(pred4),]


ggplot(pred4, aes(x = time, y = realX)) + geom_line(col="blue",size=1) + geom_line(data=pred4, aes(x=time,y=predX),col="black",size=1)

ggplot(pred4, aes(x = time, y = realY)) + geom_line(col="blue",size=1) + geom_line(data=pred4, aes(x=time,y=predY),col="black",size=1)


corX <- cor(pred4$realX,pred4$predX)
print(paste("Rsquared X=",corX^2,sep=""))

corY <- cor(pred4$realY,pred4$predY)
print(paste("Rsquared Y=",corY^2,sep=""))

pred4$residual <- sqrt((pred4$realX-pred4$predX)^2 + (pred4$realY-pred4$predY)^2)
print(paste("Overall RMSE=",sqrt(var(pred4$residual, na.rm=T)),sep=""))

distanceError <- sqrt((pred4$predX - pred4$realX)^2 + (pred4$predY - pred4$realY)^2)
mean(distanceError)
sd(distanceError)

error <- as.data.frame(distanceError)
ggplot(error, aes(x=distanceError)) + geom_density(alpha=.2, fill="#FF6666")

```




# Test walks

## A) Walk with stops through the center of the room (cross pattern)

```{r,echo=FALSE}
# Load the file
lines <- paste(readLines("locator-app-1432125956044-testwalk1.json"),collapse="")
# Retouch it a bit so that it is proper json
lines <- gsub("[]", "", lines, fixed=T)
lines <- gsub("][", ",", lines, fixed=T)
# We parse the json
json_data <- fromJSON(lines)
df2 <- ldply(json_data, data.frame)
# We add the relative timestamp since beginning of the experiment
mintime <- min(df2$timestamp)
df2$reltime <- (df2$timestamp - mintime)

# We add, at each point in time, the real position, and calculate the real distance from each beacon
realXseries2 = c(rep(1,times=10),
                seq(from=1,to=1,length.out=5),
                rep(1,times=10),
                seq(from=1,to=-1,length.out=5),
                rep(-1,times=10),
                seq(from=-1,to=2,length.out=5),
                rep(2,times=10),
                seq(from=2,to=2,length.out=5),
                rep(2,times=10),
                seq(from=2,to=0,length.out=5),
                rep(0,times=10),
                0
                )

realYseries2 = c(rep(0,times=10),
                seq(from=0,to=4.5,length.out=5),
                rep(4.5,times=10),
                seq(from=4.5,to=4.5,length.out=5),
                rep(4.5,times=10),
                seq(from=4.5,to=4.5,length.out=5),
                rep(4.5,times=10),
                seq(from=4.5,to=7,length.out=5),
                rep(7,times=10),
                seq(from=7,to=7,length.out=5),
                rep(7,times=10),
                7
                )


df2$realX = NA
df2$realY = NA
for(i in 1:length(realXseries2)){
    df2[round(df2$reltime/1000)==(i-1),"realX"] = realXseries2[i]
    df2[round(df2$reltime/1000)==(i-1),"realY"] = realYseries2[i]
}
#summary(df2)

# We reformat the data for prediction according to the model to test
topred <- data.frame(time=unique(round(df2$reltime/1000)),realX=NA,realY=NA)
for(t in topred$time){
    topred[topred$time==t,"realX"] <- unique(df2[round(df2$reltime/1000)==t,"realX"])[[1]]
    topred[topred$time==t,"realY"] <- unique(df2[round(df2$reltime/1000)==t,"realY"])[[1]]
}
for(beacon in beacons$beaconID){
    label = beacons[beacons$beaconID==beacon,"beaconNr"];
    # We create the field distance
    topred[,paste("distance",label,sep="")] <- NA
    # We go though the time values and fill in the data
    for(t in topred$time){
        topred[topred$time==t,paste("distance",label,sep="")] <- df2[round(df2$reltime/1000)==t & df2$beaconID==beacon,"distance"][1]
    }

    # We create the field rssi
    topred[,paste("rssi",label,sep="")] <- NA
    # We go though the time values and fill in the data
    for(t in topred$time){
        topred[topred$time==t,paste("rssi",label,sep="")] <- df2[round(df2$reltime/1000)==t & df2$beaconID==beacon,"rssi"][1]
    }

    # We create the field proximity
    topred[,paste("proximity",label,sep="")] <- NA
    # We go though the time values and fill in the data
    for(t in topred$time){
        topred[topred$time==t,paste("proximity",label,sep="")] <- df2[round(df2$reltime/1000)==t & df2$beaconID==beacon,"proximity"][1]
    }

}

for (i in 1:6){
    column <- topred[,paste("distance",i,sep="")]
    
    #We do filtering before smoothing the signal, taking into account the size of the room (9m longer diameter)
    column[column>ROOMSIZE]<-ROOMSIZE

    ld <- loess(column~time,topred,span=0.15)
    topred[,paste("loessdistance",i,sep="")] <- predict(ld,topred)
#     plot(topred$time,topred[,paste("distance",i,sep="")], type="l")
#     lines(topred$time,topred[,paste("loessdistance",i,sep="")],col=3)
#     cat ("Press [enter] to continue")
#     line <- readline()
}



predictionstest1 <- predict(fitloesslm2, newdata=topred)

#We plot the predictions
# count=0
# for(i in topred$time){
#     count <- count+1
#     plot(beacons$beaconX,beacons$beaconY,xlim=c(-10,11),ylim=c(-8,17))
#     
#     subset <- df2[round(df2$reltime/1000)==i,]
#     points(subset$realX,subset$realY, pch="x", cex=2)
#     for(beacon in beacons$beaconID){
#         draw.circle(beacons[beacons$beaconID==beacon,"beaconX"],beacons[beacons$beaconID==beacon,"beaconY"],radius=subset[subset$beaconID==beacon,"distance"],border="blue")
#     }
#     points(predictionstest1[count,"realX"],predictionstest1[count,"realY"], pch="o", cex=2)
#     cat ("Press [enter] to continue")
#     line <- readline()
# }

topred$predX <- predictionstest1[,1]
topred$predY <- predictionstest1[,2]

distanceError <- sqrt((topred$predX - topred$realX)^2 + (topred$predY - topred$realY)^2)
mean(distanceError)
sd(distanceError)

```

The error is awful! Let's re-train the same model now with both of the walks, and will test them with a different one:

```{r}
pred3$walk <- 1
topred$walk <- 2
newtraindata <- rbind(pred3,topred)
newfitloess <- lm(cbind(realX,realY) ~ loessdistance1*loessdistance2*loessdistance3*loessdistance4*loessdistance5*loessdistance6, data=newtraindata)
summary(newfitloess)

predictions <- predict(newfitloess, newdata=newtraindata)

#We plot the predictions
# count=0
# for(i in pred3$time){
#     count <- count+1
#     plot(beacons$beaconX,beacons$beaconY,xlim=c(-10,11),ylim=c(-8,17))
#     
#     subset <- df[round(df$reltime/1000)==i,]
#     points(subset$realX,subset$realY, pch="x", cex=2)
#     for(beacon in beacons$beaconID){
#         draw.circle(beacons[beacons$beaconID==beacon,"beaconX"],beacons[beacons$beaconID==beacon,"beaconY"],radius=subset[subset$beaconID==beacon,"distance"],border="blue")
#     }
#     points(predictions[count,"realX"],predictions[count,"realY"], pch="o", cex=2)
#     cat ("Press [enter] to continue")
#     line <- readline()
# }
# for(i in topred$time){
#     count <- count+1
#     plot(beacons$beaconX,beacons$beaconY,xlim=c(-10,11),ylim=c(-8,17))
#     
#     subset <- df2[round(df2$reltime/1000)==i,]
#     points(subset$realX,subset$realY, pch="x", cex=2)
#     for(beacon in beacons$beaconID){
#         draw.circle(beacons[beacons$beaconID==beacon,"beaconX"],beacons[beacons$beaconID==beacon,"beaconY"],radius=subset[subset$beaconID==beacon,"distance"],border="blue")
#     }
#     points(predictions[count,"realX"],predictions[count,"realY"], pch="o", cex=2)
#     cat ("Press [enter] to continue")
#     line <- readline()
# }

# Scatterplot comparison of prediction and real
ggplot(data.frame(predictions)[1:97,], aes(x = realX, y = realY)) + geom_point(alpha=0.2,col="blue",size=5) + geom_point(data=pred3,alpha=0.5,col="black",size=5)
ggplot(data.frame(predictions)[98:180,], aes(x = realX, y = realY)) + geom_point(alpha=0.2,col="blue",size=5) + geom_point(data=topred,alpha=0.5,col="black",size=5)


newtraindata$predX <- predictions[,1]
newtraindata$predY <- predictions[,2]

distanceError <- sqrt((newtraindata$predX - newtraindata$realX)^2 + (newtraindata$predY - newtraindata$realY)^2)
mean(distanceError)
sd(distanceError)
error <- as.data.frame(distanceError)
ggplot(error, aes(x=distanceError)) + geom_density(alpha=.2, fill="#FF6666")

# plot realX and predX over time
ggplot(newtraindata,aes(x=time,y=realX,color=as.factor(walk)))+geom_line()+geom_line(aes(x=time,y=predX,color=as.factor(walk),size=0.5))

# What if we re-smooth the predicted values?
# Split the data into walks before doing the smoothing!
finaldata <- data.frame()
for(i in unique(newtraindata$walk)){
    data <- newtraindata[newtraindata$walk==i,]
    loessX <- loess(predX~time,data,span=0.2)#After several tries, this seems to be the best smoothing value!
    loessY <- loess(predY~time,data,span=0.2)
    data$predX2 <- predict(loessX,newdata=data)
    data$predY2 <- predict(loessY,newdata=data)
    finaldata <- rbind(finaldata,data)
}

# We plot the new predicted values again
ggplot(finaldata,aes(x=time,y=realX,color=as.factor(walk)))+geom_line()+geom_line(aes(x=time,y=predX2,color=as.factor(walk),size=0.3))
ggplot(finaldata,aes(x=time,y=realY,color=as.factor(walk)))+geom_line()+geom_line(aes(x=time,y=predY2,color=as.factor(walk),size=0.3))
#    line <- readline()
#    cat ("Press [enter] to continue")

distanceError <- sqrt((finaldata$predX2 - finaldata$realX)^2 + (finaldata$predY2 - finaldata$realY)^2)
mean(distanceError)
sd(distanceError)
# We plot the density distribution of error
error <- as.data.frame(distanceError)
ggplot(error, aes(x=distanceError)) + geom_density(alpha=.2, fill="#FF6666")
```




## B) Continuous walk (slowish) around the room, but not the walls, changing directions

TODO: We now re-test with a more difficult walk, with less pauses